---
title: "roulette"
author: "Russ Poldrack"
date: "February 28, 2016"
output: html_document
runtime: shiny
---

This notebook simulates a simplified version of the game of roulette. The goal is to understand the various ways in which risk attitudes and value functions might make the game attractive to an individual.

First, let's simulate the outcome from 100 spins, doing this 1000 times and looking at the distribution of total gain/loss outcomes.

```{r, echo=FALSE,warnings=FALSE}
library(ggplot2)

roulette_spin = function() {
  npockets=c(18,18,2)
  roll=sample(1:sum(npockets),1)
  if (roll < npockets[1]) {
    color=0
  } else if (roll < sum(npockets[1:2])) {
    color=1
  } else {color=2}
  return(c(roll,color))
}
```

```{r, echo=FALSE,warnings=FALSE}

source("../accumulator/multiplot.R")

outcome=c()
first_win=c()

bet_value=1
bet=0  # bet on specific pocket every time - 35X payoff
payoff=35 # payoff for single pocket bet
nruns=1000
nplays=100 # how many games played at one sitting
p_payoff=1/38.

print(sprintf('Expected value of $1 bet = p(outcome)*u(outcome) = $%0.2f',-1*bet_value*(1-p_payoff) + payoff*p_payoff))

for (i in 1:nruns) {
    o=c() # spint outputs
    fw=-1 # number of spins to first win 
    for (j in 1:nplays) {
      sp=roulette_spin()
      if (payoff==sp[1] & fw<0){
        fw=j
      }
      o=rbind(o,payoff*bet_value*(payoff==sp[1]) - bet_value)
    }
    first_win=rbind(first_win,fw)
    outcome=rbind(outcome,sum(o))
}

print(sprintf('Mean outcome over %d bets (on $%0.2f bet): $%0.2f (range: $%0.2f - $%0.2f)',
      nplays,bet_value,mean(outcome),min(outcome),max(outcome)))

df=data.frame(first_win=first_win,outcome=outcome,row.names=c(1:length(outcome)))
h1=ggplot(data=df, aes(outcome)) + aes(y = ..density..) + geom_histogram(binwidth=10) + xlab(sprintf('Winnings after %d plays',nruns))
h2=ggplot(data=df, aes(first_win)) + aes(y = ..density..) + geom_histogram(binwidth=2) + xlab(sprintf('Number of plays to first win (-1:no wins)',nruns))

multiplot(h1,h2)
```



Now let's see how change in the utilty function and probability weighting function change the attractiveness of the bet. We use a prospect theory model with a power utility function and a one-parameter (Prelec) weighting function. Default values are from Table 11.3 in the Fox & Poldrack Neuroeconomics chapter

```{r, echo=FALSE,warnings=FALSE}
pi = function(p,gamma) {
  return(exp(-1*(-1*log(p))**gamma))
}

plot_value_function = function(input){
    alpha=input$alpha
    bet_value=input$bet_value
    lam=input$lambda
    payoff=35
    gamma=input$gamma
    p_payoff=1/38
    print(input)
    bet_max=1000
    beta=alpha # use same weighting for gains and losses
    o=seq(0,bet_max,1)
    v_o=o^alpha
    eu=((bet_value*payoff)^alpha)*pi(p_payoff,gamma) - lam*(bet_value^beta)
    betvalstring=sprintf('Expected value of $%d bet = $%0.2f',bet_value,eu)


    plot(o,v_o,type='l',col='blue',xlim=c(-1*bet_max,bet_max),ylim=c(-1*bet_max,bet_max),lwd=2,main=betvalstring)
    
    lines(-1*o,-1*v_o*lam,col='blue',type='l',xlim=c(-1*bet_max,bet_max),ylim=c(-1*bet_max,bet_max),lwd=2)
    lines(o,o,type='l')
    lines(-1*o,-1*o,type='l')
    lines(c(-bet_max,bet_max),c(0,0))
    points(bet_value,bet_value^alpha,col='red',pch=19,cex=2)

#     plt.axis([-1*bet_max,bet_max,-1*bet_max,bet_max])
#     plt.plot(o,o,color='black')
#     plt.plot([-1*bet_max,bet_max],[0,0])
#     plt.plot(-1*o,-1*o,color='black')
# 
#     plt.plot(bet_value,bet_value**alpha,marker='o',color='red')
# 

}

plot_weighting_function = function(gamma){
    x=seq(0,1,0.001)
    plot(x,pi(x,input$gamma),type='l',col='blue',lwd=3)
    lines(x,x,type='l')

}

```

```{r, echo=FALSE,warnings=FALSE}
inputPanel(

  sliderInput("bet_value", label = "Amount of bet",
              min = 1, max = 1000, value = 10, step = 1),
  sliderInput("lambda", label = "lambda (loss aversion coefficient)",
              min = 0.01, max = 6.0, value = 2.0, step = 0.01),
  sliderInput("alpha", label = "alpha (curvature of value function)",
              min = 0.0, max = 2.0, value = 0.88, step = 0.01),
  sliderInput("gamma", label = "gamma (curvature of probability weighting function)",
              min = 0.01, max = 0.99, value = 0.74, step = 0.01)
)

renderPlot({
  par(mfrow=c(1,2))
  plot_value_function(input)
  plot_weighting_function(input$gamma)
  })
```




