---
title: "hothand"
author: "Russ Poldrack"
date: "March 6, 2016"
output: html_document
---

This notebook provides an example of the difficulties that humans exhibit in distinguishing random from nonrandom outcomes.

First we define a function to generate a random series of free throws with a particular likelihood of success. We also provide the ability to generate a series that is streakier than random (using an autoregressive process) or one that is less streaky than random (by randomly inserting switches in the timeseries).


```{r, echo=FALSE}
mk_random=function(nshots,psuccess=0.5,ar=0.0,pswitch=0){
  
    if (ar==0 && pswitch==0) {
        r=runif(nshots)
    } else if (ar>0) {
        # generate random timeseries using autoregressive model
        r=array(dim=nshots)
        r[1]=runif(1)
        for (i in 2:nshots)
            r[i]=r[i-1]*ar + runif(1)*(1-ar)
            # make sure it's within bounds
            if (r[i]>1){
                r[i]=1
            } else if (r[i]<0) {
                r[i]=0
            }
      
    } else {
        # generate "super-random" series by flipping 
        r=array(dim=nshots)
        r[1]=runif(1)
        for (i in 2:nshots) {
            if (runif(1)<pswitch) {
                r[i]=abs(1-r[i-1])
            } else {
                r[i]=runif(1)
            }
        }
    }
  
    return(as.logical(r<psuccess))
}


```

Generate a large number of sequences, and plot the histogram of run lengths

```{r echo=FALSE}
nruns=1000
nshots=25
runlengths_rand=c()
data_rand=matrix(nrow=nruns,ncol=nshots)

for (i in 1:nruns) {
    s=mk_random(nshots)
    for (l in rle(s)$lengths) {
      runlengths_rand=c(runlengths_rand,l)
    }
    data_rand[i,]=s
}

runlengths_ar=c()
data_ar=matrix(nrow=nruns,ncol=nshots)

for (i in 1:nruns) {
    s=mk_random(nshots,ar=0.8)
    for (l in rle(s)$lengths) {
      runlengths_ar=c(runlengths_ar,l)
    }
    data_ar[i,]=s
}


runlengths_switch=c()
data_switch=matrix(nrow=nruns,ncol=nshots)

for (i in 1:nruns) {
    s=mk_random(nshots,pswitch=0.5)
    for (l in rle(s)$lengths) {
      runlengths_switch=c(runlengths_switch,l)
    }
    data_switch[i,]=s
}
      
bins=seq(0.5,max(runlengths_ar)+1.5,1)
h_rand=hist(runlengths_rand,breaks=bins,plot=FALSE)
h_ar=hist(runlengths_ar,breaks=bins,plot=FALSE)
h_switch=hist(runlengths_switch,breaks=bins,plot=FALSE)

plot(bins[1:(length(bins)-1)],h_rand$density,type='l',lwd=2,col='red',ylim=c(0,1),xlab='Run length',ylab='Probability')
lines(bins[1:(length(bins)-1)],h_ar$density,type='l',lwd=2,col='green')
lines(bins[1:(length(bins)-1)],h_switch$density,type='l',lwd=2,col='blue')
legend('topright',legend = c('random','AR','switching'),col=c('red','green','blue'),lwd=2)
```

Let's see how well we can detect runs using the standard statistical approach (runs test)

```{r echo=FALSE}
library(randtests)
library(knitr)

runstest_rand=c()

for (i in 1:nruns){
  r=runs.test(as.numeric(data_rand[i,]),threshold=0.5)
  runstest_rand=c(runstest_rand,r$p.value)
}

runstest_ar=c()

for (i in 1:nruns){
  r=runs.test(as.numeric(data_ar[i,]),threshold=0.5)
  runstest_ar=c(runstest_ar,r$p.value)
}

runstest_switch=c()

for (i in 1:nruns){
  r=runs.test(as.numeric(data_switch[i,]),threshold=0.5)
  runstest_switch=c(runstest_switch,r$p.value)
}
knit_print('Probability of rejecting null hypothesis of random generation')
sigp=c(mean(runstest_rand<0.05,na.rm=TRUE),
       mean(runstest_ar<0.05,na.rm=TRUE),
       mean(runstest_switch<0.05,na.rm=TRUE))
sigp_df=as.data.frame(sigp,row.names=c('random','AR','switching'))
kable(sigp_df)
```

Now let's use the conditional probability approach (discussed here: http://andrewgelman.com/2015/07/09/hey-guess-what-there-really-is-a-hot-hand/)

```{r, echo=FALSE}
probtest_rand=c()

for (i in 1:nruns){
  h1=data_rand[i,1:(nshots-1)]
  h2=data_rand[i,2:nshots]
  probtest_rand=c(probtest_rand,sum(h1 & h2)/sum(h1))
}

probtest_ar=c()

for (i in 1:nruns){
  h1=data_ar[i,1:(nshots-1)]
  h2=data_ar[i,2:nshots]
  probtest_ar=c(probtest_ar,sum(h1 & h2)/sum(h1))
}

probtest_switch=c()

for (i in 1:nruns){
  h1=data_switch[i,1:(nshots-1)]
  h2=data_switch[i,2:nshots]
  probtest_switch=c(probtest_switch,sum(h1 & h2)/sum(h1))
}

estimates=c(mean(probtest_rand), mean(probtest_ar,na.rm=TRUE),mean(probtest_switch))
est_df=as.data.frame(estimates,row.names=c('random','AR','switching'))
kable(est_df)


```