---
title: "SPRT example"
author: "Russ Poldrack"
date: "February 27, 2016"
output: html_document
runtime: shiny
---

This code generates an example of Wald's sequential probability ratio test (SPRT). In this example, we will decide whether a noisy signal is moving either up or down.  First we need to define functions to compute the log likelihood ratio of two Gaussians, to generate data from a diffusion process, and to fit the SPRT to the data.  We will arbitrarily define 1 and -1 as the centers of the distributions for up and down motion.

*NOTE:* There will be a relatively long delay initially as well as after you change the slider settings, because the simulation must be re-run each time.

```{r, echo = FALSE,warnings=FALSE}
require(ggplot2)
source('multiplot.R')

gauss_LLR <- function(xbar, sigma=1,mu=c(1,-1)) {
return(log(dnorm(xbar,mean=mu[1],sd=sigma)/dnorm(xbar,mean=mu[2],sd=sigma)))
}

mkdata <- function(drift=0.01,noise_sd=0.1,npts=1000) {
  
  d=array(data=NA,dim=npts)
  d[1]=rnorm(1)*noise_sd
  for (i in 2:npts){
    d[i]=drift+rnorm(1)*noise_sd
  }
  return(d)
}

cumulate <- function(d) {
  cumul=array(NA,dim=length(d))
  cumul[1]=d[1]
  for (i in 2:length(d)){
    cumul[i]=sum(d[1:i])
  }
 return(cumul) 
}


fit_sprt <- function(d,alpha=0.05) {
  # define acceptable specificity and sensitivity
   # alpha is likelihood of deciding A when A is true
  beta=1.-alpha # likelihood of deciding B when B is true
  
  A=log(beta/alpha)
  B=log((1.-beta)/(1.-alpha))
  
  npts=length(d)
  ll=c()
  outcome=NA
  for (i in 1:npts) {
    ll=rbind(ll,gauss_LLR(d[i]))
    S=sum(ll)
    if (S>A) {
      outcome=1
      break
    } else if (S<B) {
      outcome=-1
      break
    }
  }
  return(c(outcome,i))
}

        
```


Now let's generate a number of trials and examine the trajectories along the distribution of finishing times.
```{r, echo = FALSE,warnings=FALSE}

plot_sprt = function(drift) {
  nruns=1000
  npts=500
  sprt=matrix(data=NA,nrow=nruns,ncol=2)
  cumul=matrix(data=NA,nrow=nruns,ncol=npts)
  
  for (r in 1:nruns) {
      d=mkdata(npts=npts,drift=drift)
      sprt[r,]=fit_sprt(d)
      cumul_data=cumulate(d)[1:sprt[r,2]]
      cumul[r,1:sprt[r,2]]=cumul_data
  
  }
  
  df=data.frame(t(cumul))
  df$index=c(1:npts)
  
  sprt_df=data.frame(resp=as.factor(sprt[,1]),rt=sprt[,2])
  sprt_df_finished=sprt_df[!is.na(sprt_df$resp),]  # drop empty rows
  h=ggplot(data=sprt_df_finished, aes(rt,colour=resp,fill=resp)) + geom_density(size=2,adjust=1/2,alpha=0.5) +  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.title.x=element_blank(),legend.position='top') + xlim(0,npts)
  
  
  yrange=max(abs(range(cumul,na.rm=TRUE)))
  nruns_to_show=min(20,nruns)
  for (r in 1:nruns_to_show) {
      goodpts=!is.na(cumul[r,])
      tmpdf=data.frame(xdata=df$index[goodpts],ydata=cumul[r,goodpts])
      if (r==1){ 
          p=ggplot(data=tmpdf,aes(x=xdata,y=ydata))+geom_line() + xlim(0,npts)+  theme(axis.text.y=element_blank())+ylab('evidence')+xlab('timepoints')+ggtitle(sprintf('mean rt = %0.1f, accuracy=%0.3f, nonresponse=%0.3f',mean(sprt[sprt[,1]==1,2],na.rm=TRUE),mean(sprt[,1]==1,na.rm=TRUE),mean(is.na(sprt[,1]))))
      } else {
          p=p+geom_line(data=tmpdf,aes(x=xdata,y=ydata))
          #lines(cumul[r,])
      }
  }
  multiplot(h, p)
}
```

```{r, echo = FALSE,warnings=FALSE}

inputPanel(
      sliderInput("drift_rate", "drift rate:",
                  min=0.0005, max=0.03, value=0.005,step=0.0005))
renderPlot({
  plot_sprt(input$drift_rate)
  })


```